---
title: "R Notebook"
output: html_notebook
---

Using the r package `overlap` following the approach described in Gaynor et al. 2018, Effects of human settlement and roads on diel activity patterns of elephants (Loxodonta africana). African Journal of Ecology 56:872-881.

Set things up:

```{r}
rm(list = ls()) #clear's R's brain - removes objects from environment
library(tidyverse)
#install.packages("overlap")
library(overlap)
library(lubridate)
```

Pull in the data

```{r}
DF <- read.csv("./Data/Activity.csv")

summary(DF$Hour)
```
Now convert hours to radians

```{r}
DF<- DF %>% mutate(
   Hours_rad = Hour*2*pi
)


```

Now let's do a density plot!
```{r}
#first extract data as vector
Time <- DF$Hours_rad
densityPlot(Time)
```
Threw an error because of NA's - fix that

```{r}
Time <- na.omit(Time)
densityPlot(Time)
```
Fix error of numeric vector.

```{r}
Time <- as.numeric(Time)
densityPlot(Time)
```
R doesn't like how we chunked the time - now divide by 24 to fix radians.

```{r}
Time<- Time/24
densityPlot(Time)
```
Summarize data set for number of observations

```{r}
Patterns<-DF %>% group_by(bin) %>% summarize(
   numObs = length(bin)
)
```
I can see that we have plenty of observations of fishers and porcupines so let's extract them

Extract porcupine and fisher data
```{r}
Porcs<-filter(DF, bin == "PORCUPINE")
Fisher <- filter(DF, bin == "FISHER")
```
And now extract the times and make them vectors

```{r}
PorcT<- Porcs$Hour
FisherT<-Fisher$Hour
```
Now convert to radians

```{r}
PorcRads <- as.numeric((PorcT/24)*2*pi)
FisherRads <- as.numeric((FisherT/24)*2*pi)

densityPlot(PorcRads)
densityPlot(FisherRads)
```

Create overlap plot
```{r}
overlapPlot(PorcRads, FisherRads)

#now a prettier version

MyPlot<-overlapPlot(PorcRads, FisherRads, linet = c(1,1), linec = c("red", "blue"),
  rug=TRUE, extend="lightgreen", main="Porcupine vs. Fisher activity")
legend("topleft", c("Porcupine", "Fisher"), lty=1, col=c("red", "blue"), bg="white")
MyPlot
```
Now calcualte d-hat:

```{r}
overlapEst(PorcRads, FisherRads )
```
NOw get confidence intervals:

First do 999 smoothed bootstrap values for the pair of species:

```{r}
Dhat1<- overlapEst(PorcRads, FisherRads, type = "Dhat1")
boot <- bootstrap(PorcRads, FisherRads, 999, type = "Dhat1")
mean(boot)
hist(boot)
abline(v=Dhat1, col = "red", lwd = 2)
abline(v = mean(boot), col = "blue", lwd = 2, lty = 3)
```
Now calculate confidence intervals:

```{r}
#95% CI is default
bootCI(Dhat1, boot) ['norm0',]
bootCI(Dhat1, boot) ['basic0',]

#now do a 90% confidence interval

bootCI(Dhat1, boot, conf = 0.9)['norm0',]
```








